El Mayor Postseismic \author{Trever Hines} \rule{\headwidth}{1.0pt}

### 2016-03-29 RoughDraft/ElMayor.tex
commit before push

\section*{Introduction}

@@

Given the inability to describe both near and far field deformation with fault slip in an elastic lithosphere, \citet{Pollitz2012}, \citet{Rollins2015} and \citet{Spinler2015} have explored viscoelastic relaxation in the lower crust and upper mantle as a potential deformation mechanism. The lithospheric rheology is largely unknown and so modeling postseismic deformation with viscoelastic relaxation requires one to assume a rheologic model and find the best fitting model parameters, which is generally a computationally expensive nonlinear inverse problem. Consequently, a simplified structure for the lithosphere must be made to minimize the number of rheologic parameters that need to be estimated.  For example, it is commonly assumed that the lithosphere consists of only three homogeneously Maxwell viscoelastic layers, which may be an inadequate representation of the lithosphere \citep[e.g.][]{Hines2013,Riva2009}. Additionally, it is necessary to make simplifying assumptions about the nature of afterslip. For example, one can assume a frictional model for afterslip and parameterize afterslip in terms its the unkown rheologic properties \citep[e.g.][]{Johnson2009,Johnson2004}. Additionally, one can assume that afterslip only dominates the first year of deformation \citep[e.g.][]{Pollitz2012,Spinler2015} and that viscoelastic relaxation is the dominant mechanism in later years. However, numerous observations of interseismic fault creep has been recognized in this and similar tectonic settings and it has been speculated that such creep can be initiated as a postseismic process \citep{Cakir2012,Cetin2014}.  We therefore do not make to implicit assumption that afterslip terminates after a given amount of time.  Indeed, the preferred viscoelastic model from \citet{Pollitz2012} underestimates near field velocities, which could be indicative of unmodeled continued afterslip.

@@

\section*{Data Processing}

We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) stations within a 400 km radius about the El Mayor-Cucapah epicenter. Our analysis is on the coseismic and postseismic deformation resulting from the El Mayor-Cucapah earthquake, which we collectively describe as $u(t)$. We consider GPS position time series, $u_\mathrm{obs}(t)$, to be the superposition of $u(t)$, secular tectonic deformation, annual and semi-annual fluctuations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010 Mw5.8 Ocotillo earthquake and the August 26, 2012 Brawley swarm, which consisted of a Mw5.5 and Mw5.3 event, are the only earthquakes after the EMC earthquake that produced noticeable offsets recorded by GPS. Although the Ocotillo earthquake had its own series of aftershocks (Haukson), neither earthquake produced transient deformation that is detectable with GPS. We thus model $u_\mathrm{obs}(t)$ as

\begin{equation}
  u_\mathrm{obs}(t) = u_\mathrm{pred}(t) + \epsilon
\end{equation}
where
\begin{equation}\label{TimeSeriesModel}
  \begin{split}  
    u_\mathrm{pred}(t) = &u(t) + c_0 + c_1t + \\
                         &c_2\sin(2\pi t) + c_3\cos(2\pi t) + c_4\sin(4\pi t) + c_5\cos(4\pi t) + \\
                         &c_6H(t-t_\mathrm{oc}) + c_7H(t-t_\mathrm{bs}).
  \end{split}
\end{equation}
In the above equations, $t_\mathrm{oc}$ and $t_\mathrm{bs}$ are the times of the Ocotillo earthquake and Brawley swarm respectively, $H(t)$ is the Heaviside function, $c_0$ through $c_7$ are unknown coefficients, and $\epsilon$ is noise with zero mean and variance that assumed known.

Stations which recorded signals that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, which record deformation that is largely anthropogenic. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. While several stations were installed after the EMC earthquake to improve the spatial resolution of postseismic deformation \citep{Spinler2015}, our inverse method uses postseismic displacements rather than velocities (e.g. ...)
ELABORATE ON VELOCITY VS DISPLACEMENT
, which requires the knowledge of the stations preseismic position. Despite our inability to utilize potentially rich data, we prefer to use postseismic displacements rather than potentially dubious estimates of postseismic velocities.

The October 16, 1999 Hector Mine earthquake, which occurred within our study region about 270 km north of the EMC epicenter, has produce transient postseismic deformation which we do not wish to model either mechanically or through empirical line fitting. We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, past which point postseismic deformation for nearfield sites occurs at an approximately steady rate \citep{Savage2009}. When considering stations further away from the Hector Mine epicenter, postseismic transience persists for only about two years \citep{Spinler2015}.

Studies of postseismic deformation typically assume a parametric form for $u(t)$ such as such as a logarithmic or exponential function. Such functions have been found empirically e.g. \citep{Savage2005a} and theoretically \citep{Marone1991} to adequately describe deformation resulting from postseismic deformation. However, by assuming a logarithmic or exponential form of $u(t)$ we run the risk of overfitting the GPS time series and inferring a nonexistent postseismic signal. We therefore do not assume any parameteric form for $u(t)$ and rather treat it as integrated Brownian motion, so that 
\begin{equation}
    \dot{u}(t) = \sigma^2\int_0^t w(s) ds.
\end{equation}

where $w(t)$ is white noise and the variance of $\dot{u}(t)$ increases linearly with time by a factor of $\sigma^2$. We use a
Kalman filtering approach to estimate $u(t)$ and the unknown parameters in eq. \ref{TimeSeriesModel} which we describe now.  In the context of Kalman filtering, our time varying state vector is
\begin{equation}
    \mathbf{X}(t) = [u(t),\dot u(t), c_0, ..., c_7]
\end{equation}
and eq. \ref{TimeSeriesModel} is the observation function which maps the state vector to the GPS observations. We initiate the Kalman filter by assuming a prior estimate of $\mathbf{X}(t)$ at time $t_0$ which has a sufficiently large covariance to effectively make our prior uninformed.  For each time epoch, $t_i$, Bayesian linear regression is used to incorperate GPS derived estimates of displacement with our prior estimate of the state, $\mathbf{X}_{i|i-1}$, to form a postserior estimate of the state, $\mathbf{X}_{i|i}$, which has covariance $\mathbf{\Sigma}_{i|i}$.

@@

This process is repeated for each of the $N$ time epochs at which point we use Rauch-Tung-Striebel smoothing to find $X_{i|N}$, which is an estimate of the state at time $t_i$ that incorporates all $N$ GPS observation.  Our final estimates of $u(t)$ are used in subsequent analysis, while the remaining components of of the state vector are considered nuisance parameters. In the interests of computational tractability, we down sample our smoothed time series from daily solutions down to weekly solutions.

The smoothness of $u(t)$ is controlled by the chosen value of $\sigma^2$, which describes how rapidly we expect the postseismic signal to vary over time.  Setting $\sigma^2$ equal to zero will effectively result in modeling $u(t)$ as a straight line which is insufficient to describe the expected transient behavior in postseismic deformation \citep{Savage2005a}. The other endmember, where $\sigma^2$ is infinitely large, will result in $u_\mathrm{pred}(t)$ fitting what is obviously noise in the data. While one can use a maximum likelihood based approach to picking $\sigma^2$ (e.g. \citep{Segall1997}), we rather take a subjective approach and choose a value for $\sigma^2$, which is used when filtering time series for each station, that is just large enough to faithfully describes the early rapid rate of postseismic deformation at the most near field station in our study, P496.  This ensures that $\sigma^2$ will be sufficiently large so that our estimate of $u(t)$ does not smooth out potentially valuable postseismic signal. We find that when using $\sigma^2 = 0.05 \mathrm{m}^2 / \mathrm{yr}^3$, we are able to adequately describe all but the first week of postseismic deformation at station P496, which gets incorporated into our estimate of coseismic displacements (figure \ref{P496Fit}).  Since we assume that the first week of deformation is overwhelmingly the result of afterslip, any unmodeled afterslip over the first week following the El Mayor-Cucapah earthquake will be added into our estimates of coseismic slip in section ().  Figure \ref{P496PS} shows the estimate of coseismic and postseismic deformation, $u(t)$, for station P496 along with its formal uncertainties.

It is important to note that the shown uncertainties in $u(t)$ do not account for the non-negligible epistemic uncertainty in eq. \ref{TimeSeriesModel}.  For example, we assume that the background rate of deformation can be approximated as having a constant rate, while mechanical models would suggest this to not be true on earthquake cycle timescales \citep{Thatcher1983}. Although based on inspection,  assuming a constant rate of background deformation appears to be appropriate for all stations except for perhaps the stations closest to the Hector Mine epicenter, where postseismic deformation persists.  Also, our model for seasonal deformation in eq. \ref{TimeSeriesModel} assumes a constant amplitude over time, which means that years of particularly heavy or light rainfall cannot be adequately described.  This deficiency in our seasonal model causes our estimate of $u(t)$ to describe some of the unmodeled oscillations (figure \ref{P619PS}).

note that stations jumps are only estimated for stations within 40 km of BS or OC

Cite Freed 2007 for far reaching postseismic after Landers/Hector Mine.

We show the difference in $u(t)$ for each station from 0 to 0.8 years (figure \ref{nearfield2} and \ref{farfield2}), 0.8 to 3.2 years (figure \ref{nearfield3} and \ref{farfield3}), and from 3.2 to 5.0 years (figure \ref{nearfield4} and \ref{farfield4}).  For far field stations (figure \ref{farfield2}, \ref{farfield3}, and \ref{farfield4}), we can see a clear south trending displacement throughout the first 3.2 years in stations as far as 400 km from of the El Mayor Cucapah epicenter.  These displacements are most pronounced along the direction of the El Mayor-Cucapah P and S axis, as would be expected for postseismic deformation.  Beyond 3.2 years, The southward trending far field postseismic deformation is barely perceptible. There is however, an eastward trend in far field stations just west of the Hector Mine epicenter.  We suspect this this deformation, which is inconsistent with the pattern expected for El Mayor-Postseismic, is the result of continued transient postseismic deformation following the Hector Mine earthquake.  The vertical deformation in the far field is difficult to attribute to postseismic processes.  Most far field stations display an initial subsidence for the first year after the El Mayor-Cucapah earthquake followed by continued uplift.  This trend in vertical deformation can be observed in all three of the quadrants where postseismic data is available, which means that the vertical deformation does not exhibit an anti-symetric quadrant pattern, as would be expected for postseismic processes.  Although we use vertical deformation in our subsequent analysis, we do not put an emphasis on trying to describe the vertical deformation as it likely has non-tectonic origins.

Near field deformation is show in figures \ref{nearfield2}, \ref{nearfield3}, and \ref{nearfield4}.  The near field deformation is notably sustained when compared to the far field deformation.  Namely, the station in this study which is closest to the El Mayor-Cucapah epicenter, P496, is moving at a steady rate of ~1.4 cm/yr to the south five years after the El Mayor-Cucapah earthquake.  Vertical postseismic deformation in the near field does display a quadrant pattern which is consistent with the coseismic vertical deformation, suggesting that it is indeed resulting from tectonic processes.  However, the vertical postseismic deformation signal is only apparent in the early postseismic period (figure \ref{nearfield2}).  As with the far field deformation, there is a general trend of uplift in the near field one year after the earthquake which we do not consider to be related to postseismic processes.

\section*{Postseismic Modeling}

In this paper, we seek to find the mechanisms driving five years of postseismic deformation following the El Mayor-Cucapah earthquake. We consider afterslip and viscoelastic relaxation in the lithosphere and/or asthenosphere as candidate mechanisms.  Poroelastic rebound has also been used to model postseismic deformation \cite{Jonsson2003}; however, \cite{Gonzalez-ortega2014} suggest that any contribution to postseismic deformation from poroelastic rebound would be negligible.  Additionally, we are considering stations which are sufficiently far away from the main rupture that poroelastic rebound should be insignificant.

\subsection*{model design}
We estimate coseismic and time dependent postseismic fault slip, both of which are assumed to occur on the fault geometry estimated by \cite{Wei2011} with some modifications.  Field studies \cite{Fletcher2014} and LIDAR observations \cite{Oskin2012} have revealed a significantly more complicated fault geometry than what was inferred by \cite{Wei2011}, especially within the Sierra Cucapah.  However, we find the relatively simple coseismic fault geometry from \cite{Wei2011a} to be sufficient because most of the stations used in this study are sufficiently far from the El Mayor-Cucapah rupture zone that they are insensitive to the details in the fault geometry found by \cite{Fletcher2014} and \cite{Oskin2012}.  The fault geometry used in this study (figure \ref{fig:ContextMap}) consists of the two main fault segments inferred by \cite{Wei2011}, where the northern segment runs through the Sierra Cucapah up to the US-Mexico border and the southern segment is the Indiviso fault which extends down to the Gulf of California.  Additionally, We extend the northern segment by 40 km to the northwest which is motivated by the clustering of aftershocks on the northern tip of coseismic rupture zone \cite{Kroll2013} \cite{Hauksson2011}.  This extended fault segment was also found to be necessary by \cite{Rollins2015} and \cite{Pollitz2012} for describing postseismic deformation.

We consider a variety of rheologic models for the lithosphere and asthenosphere in this study.  The simplest rheologic model is to consider the lithospere and asthenosphere to be effectively elastic and isotropic.  In such case, the rheologic parameters, consisting of the Lame parameters, $\lambda$ and $\mu$, and density, $\rho$, are reasonably well known. The only unknowns are then the distribution of fault slip, which can be easily estimated due to the linearity of the forward problem.  \cite{Rollins2015} found that postseismic deformation following the El Mayor-Cucapah earthquake can be well explained with afterslip on the coseismicly ruptured fault, which has been extended down to the lithosphere-asthenosphere boundary inferred by \cite{Lekic2011}, and they did not require any viscoelastic relaxation to describe the observations.  They did however, find that an unrealistically large amount of afterslip is required on the Indiviso fault, equivalent to a Mw7.2 earthquake.  Since most of the GPS data is located north of the rupture zone, slip on the Indiviso fault segment is not well constrained and it could be acting as a proxy for distributed relaxation in the lower crust or upper mantle.  It is well understood that afterslip at sufficiently great depths can explain postseismic deformation resulting from viscoelastic relaxation \cite{Savage1990}, at least in 2D earthquake models.  In the interest of eliminating null space in our inverse problem,  we assuming afterslip occurs only at seismogenic depths ($<15$ km), and we allow for viscoelastic relaxation in the lower crust (15-30 km) and the upper mantle.  We choose such shallow constraints on afterslip because rate-state friction based models would suggest afterslip to occur at shallow depths to accommodate a coseismic slip deficit \cite{Marone1991}, which is indeed inferred in the coseismic models by \cite{Wei2011a} and (Fialko 2010).  Additionally, as noted by \cite{Rollins2015} the postseismic uplift just north of the rupture zone can be well described with slip at seismogenic depths, while afterslip which is constrained to depths below 15 km predicts subsidence in that region.

Another reason for constraining slip to be shallower than 15 km depth is because we seek to estimate a lower crustal viscosity and the inverse problem becomes particularly ill-posed when trying to simultaneously estimate afterslip and a lithospheric viscosity at the same depths. 
This ill-posedness is illustrated in figure (\ref{fig:LowerCrust}), which shows the displacement resulting from a meter of slip on a fault extending from 15 to 30 km depth as well as the initial velocity resulting from viscoelastic relaxation in the lower crust, which is given a viscosity of $10^{18}$ Pa s.  The horizontal displacement from fault slip is in the opposite direction as the displacements resulting from subsequent viscoelastic relaxation.  This means that displacement resulting from afterslip at lower crustal depths can be cancelled out by a low viscosity lower crust.  We eliminate this null space by allowing only one mechanism in the lower crust, which we choose to be viscoelastic relaxation.  Aside from the issue of nonuniqueness, it is also nonsensical to allow for both brittle and ductile deformation mechanisms at the same depth.

A schematic representation of the viscoelastic rheologic models considered in this study for the lower crust and upper mantle are shown in figure (TBD).  We consider Maxwell viscoelasticity, where the steady state viscosity $\eta_M$ is unknown; Zener viscoelasticty, where the transient viscosity $\eta_K$ and transient shear modulus $\mu_K$ are unknown; and Burgers viscoelasticity, where $\eta_M$, $\eta_K$, and $\mu_K$, are all largely unknown. We further discuss these rheologic models and their use in geophysical studies in the discussion.

\subsection*{constraint on effective viscosity}\label{InitialInversion}
For any linear viscoelastic rheology of the lithosphere and asthenosphere, postseismic displacement resulting from time dependent fault slip, $s(\xi,t)$, on a fault $F$ can be described as  
\begin{equation}\label{GeneralForward}
  u(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F s(\tau,\xi) f(t-\tau,x,\xi) d\xi d\tau
\end{equation}
where $g(x,\xi)$ describes the instantaneous elastic displacements at $x$ resulting from a unit of slip at $\xi$, and $f(t,x,\xi)$ describes the velocity at $x$ resulting from viscoelastic relaxation of stresses induced by slip at $\xi$. $f$ is a function of the properties of the chosen rheologic model, which are generally not well known.  The parameters controlling the instantaneous elastic behaviour of the lithosphere and asthenosphere, $\mu$, $\lambda$, and density $\rho$, are relatively well known and we use the values listed in table (TBD), which are consistent with the properties used in the coseismic model by \cite{Wei2011} and the 1D SCEC background velocity model.  The remaining rheologic parameter, which we will assume to have a depth dependence, are estimated to optimize the predicted fit to observable postseismic displacement.

>>>>

\section{Introduction}

@@

Given the inability to describe both near and far field deformation with fault slip in an elastic lithosphere, \citet{Pollitz2012}, \citet{Rollins2015} and \citet{Spinler2015} have explored viscoelastic relaxation in the lower crust and upper mantle as a potential deformation mechanism. The lithospheric rheology is largely unknown and so modeling postseismic deformation with viscoelastic relaxation requires one to assume a rheologic model and find the best fitting model parameters, which is generally a computationally expensive nonlinear inverse problem. Consequently, a simplified structure for the lithosphere must be made to minimize the number of rheologic parameters that need to be estimated.  For example, it is commonly assumed that the lithosphere consists of only three homogeneously Maxwell viscoelastic layers, which may be an inadequate representation of the lithosphere \citep[e.g.][]{Hines2013,Riva2009}. Additionally, it is necessary to make simplifying assumptions about the nature of afterslip. For example, one can assume a frictional model for afterslip and parameterize afterslip in terms its the unknown rheologic properties \citep[e.g.][]{Johnson2009,Johnson2004}. Additionally, one can assume that afterslip only dominates the first year of deformation \citep[e.g.][]{Pollitz2012,Spinler2015} and that viscoelastic relaxation is the dominant mechanism in later years. However, numerous observations of interseismic fault creep has been recognized in this and similar tectonic settings and it has been speculated that such creep can be initiated as a postseismic process \citep{Cakir2012,Cetin2014}.  We therefore do not make to implicit assumption that afterslip terminates after a given amount of time.  Indeed, the preferred viscoelastic model from \citet{Pollitz2012} underestimates near field velocities, which could be indicative of unmodeled continued afterslip.

@@

\section{Data Processing}\label{sec:Data}

We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) stations within a 400 km radius about the El Mayor-Cucapah epicenter. Our analysis is on the coseismic and postseismic displacement resulting from the El Mayor-Cucapah earthquake, which we collectively describe as $u(t)$. We consider GPS position time series, $u_\mathrm{obs}(t)$, to be the superposition of $u(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010 Mw5.8 Ocotillo earthquake and the August 26, 2012 Brawley swarm, which consisted of a Mw5.5 and Mw5.3 event (figure \ref{fig:ContextMap}), are the only earthquakes after the El Mayor-Cucapah earthquake that produced noticeable offsets recorded by GPS. Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither earthquake produced detectable postseismic deformation of its own. We thus model $u_\mathrm{obs}(t)$ as 
\begin{equation}
  u_\mathrm{obs}(t) = u_\mathrm{pred}(t) + \epsilon
\end{equation}
where
\begin{equation}\label{TimeSeriesModel}
  \begin{split}  
    u_\mathrm{pred}(t) = &u(t)H(t-t_\mathrm{emc}) + c_0 + c_1t + \\
                         &c_2\sin(2\pi t) + c_3\cos(2\pi t) + c_4\sin(4\pi t) + c_5\cos(4\pi t) + \\
                         &c_6H(t-t_\mathrm{oc}) + c_7H(t-t_\mathrm{bs}).
  \end{split}
\end{equation}
In the above equations, $t_\mathrm{emc}$, $t_\mathrm{oc}$ and $t_\mathrm{bs}$ are the times of the El Mayor-Cucapah earthuqake, Ocotillo earthquake, and the Brawley swarm respectively, $H(t)$ is the Heaviside function, $c_0$ through $c_7$ are unknown coefficients, and $\epsilon$ is the observation noise. We only estimate jumps associated with the Ocotillo earthquake and Brawley swarm for stations within 40 km of their epicenters.

Stations which recorded signals that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, which record deformation that is largely anthropogenic. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. Several GPS stations were installed after the El Mayor-Cucapah earthquake to improve the spatial resolution of postseismic deformation \citep{Spinler2015} and it is possible to subtract secular velocities derived from elastic block models \citep[e.g.][]{Meade2005} from velocities recorded at the newly installed stations to get an estimate of postseismic velocities. However, we use coseismic and postseismic displacements, rather than velocities, in our inverse method described in section \ref{sec:Model} because estimating velocities from an already noisy displacement time series can introduce significant aleatoric and epistemic uncertanties depending on exactly how the estimation is done. This modeling choice prevents us from using the newly installed stations in Baja California in our analysis.

The October 16, 1999 Hector Mine earthquake, which occurred within our study region about 270 km north of the El Mayor-Cucapah epicenter, has produce transient postseismic deformation which we do not wish to model either mechanically or through empirical line fitting. We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, past which point postseismic velocities at sites proximal to the Hector Mine epicenter are approximately constant \citep{Savage2009}.  When appraising our model fit in section \ref{sec:Model}, we see some systematic residuals in the vicinity of the Hector Mine, which may be the result of errors in this approximation.

Studies of postseismic deformation typically assume a parametric form for $u(t)$, such as one with a logarithmic or exponential time dependence \citep[e.g.][]{Savage2005a}.  However, by assuming a logarithmic or exponential form of $u(t)$ we run the risk of over fitting the GPS time series and inferring a nonexistent postseismic signal. We therefore do not assume any parametric form for $u(t)$ and rather treat it as integrated Brownian motion, so that 
\begin{equation}
    \dot{u}(t) = \sigma^2\int_0^t w(s) ds.
\end{equation}    
where $w(t)$ is white noise and the variance of $\dot{u}(t)$ increases linearly with time by a factor of $\sigma^2$. We use a Kalman filtering approach to estimate $u(t)$ and the unknown parameters in eq. \ref{TimeSeriesModel}.  In the context of Kalman filtering, our time varying state vector is
\begin{equation}
    \mathbf{X}(t) = [u(t),\dot u(t), c_0, ..., c_7]
\end{equation}
and eq. \ref{TimeSeriesModel} is the observation function which maps the state vector to the GPS observations. We initiate the Kalman filter by assuming a prior estimate of $\mathbf{X}(t)$ at the first time epoch, denoted $\mathbf{X}_{1|0}$, which has a sufficiently large covariance, denoted $\mathbf{\Sigma}_{1|0}$, to effectively make our prior uninformed.  For each time epoch, $t_i$, Bayesian linear regression is used to incorporate GPS derived estimates of displacement with our prior estimate of the state, $\mathbf{X}_{i|i-1}$, to form a posterior estimate of the state, $\mathbf{X}_{i|i}$, which has covariance $\mathbf{\Sigma}_{i|i}$.

@@

This process is repeated for each of the $N$ time epochs at which point we use Rauch-Tung-Striebel smoothing \citep{Rauch1965} to find $\mathbf{X}_{i|N}$, which is an estimate of the state at time $t_i$ that incorporates GPS observation for all $N$ time epochs.  Our final estimates of $u(t)$ are used in subsequent analysis, while the remaining components of the state vector are considered nuisance parameters. In the interests of computational tractability, we down sample our smoothed time series from daily solutions down to weekly solutions.

The smoothness of $u(t)$ is controlled by the chosen value of $\sigma^2$, which describes how rapidly we expect the postseismic signal to vary over time.  Setting $\sigma^2$ equal to zero will effectively result in modeling $u(t)$ as a straight line which is insufficient to describe the expected transient behaviour in postseismic deformation. The other end member, where $\sigma^2$ is infinitely large, will result in $u_\mathrm{pred}(t)$ fitting what is obviously noise in the data. While one can use a maximum likelihood based approach to picking $\sigma^2$ \citep[e.g.][]{Segall1997}, we rather take a subjective approach and choose a value for $\sigma^2$ that is just large enough to faithfully describe the postseismic deformation at the most near field station in our study, P496, which exhibits the most pronounced rapid changes in velocity. This ensures that $\sigma^2$ will be sufficiently large so that our estimate of $u(t)$ does not smooth out potentially valuable postseismic signal at the remaining stations. We find that when using $\sigma^2 = 0.05 \mathrm{m}^2 / \mathrm{yr}^3$, we are able to adequately describe all but the first week of postseismic deformation at station P496, which gets incorporated into our estimate of coseismic displacements (figure \ref{fig:P496Fit}). We assume that the first week of deformation is overwhelmingly the result of afterslip and any unmodeled afterslip over the first week following the El Mayor-Cucapah earthquake will be added into our estimates of coseismic slip in section \ref{sec:Model}.  Figure \ref{fig:P496PS} shows the estimate of coseismic and postseismic deformation, $u(t)$, for station P496 along with the estimated uncertainties. \cite{Freed2007a} noted that postseismic deformation can be observed at distances greater than 200 km from the Hector Mine earthquake and indeed, after processing the time series for stations up to 400 km from the El Mayor-Cucapah epicenter, we see clear postseismic transient deformation (\ref{fig:P619Fit} and \ref{fig:P619PS}).

It is important to note that the shown uncertainties in $u(t)$ do not account for the non-negligible epistemic uncertainty in eq. \ref{TimeSeriesModel}.  For example, we assume a constant secular rate of deformation, while viscoelastic mechanical models of the earthquake cycle predict that the accumulation of tectonic deformation is nonlinear over time \citep{Thatcher1983}. Based on inspection, the rate of secular deformation does appear constant for all stations except for perhaps the stations closest to the Hector Mine epicenter, as noted aboved.  Also, our model for seasonal deformation in eq. \ref{TimeSeriesModel} assumes a constant amplitude over time, which means that years of particularly heavy or light rainfall cannot be adequately described.  This deficiency in our seasonal model causes our estimate of $u(t)$ to describe some of the unmodeled oscillations (e.g. figure \ref{fig:P619PS}).

We show the near and far field postseismic deformation accumulated over the intervals  0-0.8 years, 0.8-3.0 years, and 3.0-5.0 years, as well as coseismic displacements in figures \ref{fig:NearField1} through \ref{fig:FarField4}.  In the far field, we can see clear south trending displacement for the first 3.0 years following the earthquake at stations as far as 400 km from of the El Mayor Cucapah epicenter.  These displacements are most pronounced along the direction of the El Mayor-Cucapah P and T axis, as would be expected for postseismic deformation. After 3.0 years, The southward trending far field postseismic deformation is barely perceptible.  The vertical deformation in the far field is difficult to attribute to postseismic processes.  Most far field stations display an initial subsidence for the first year after the El Mayor-Cucapah earthquake followed by continued uplift.  This trend in vertical deformation can be observed in all three of the quadrants where postseismic data is available, which means that the vertical deformation does not exhibit an anti-symetric quadrant pattern, as would be expected for postseismic processes.  Although we use vertical deformation in our analysis in section \ref{sec:Model},  we do not put an emphasis on trying to describe the vertical deformation as it likely has non-tectonic origins.

The near field postseismic deformation is notably sustained when compared to the far field deformation.  Namely, the station in this study which is closest to the El Mayor-Cucapah epicenter, P496, is moving at a steady rate of ~1.4 cm/yr to the south five years after the El Mayor-Cucapah earthquake.  Vertical postseismic deformation in the near field does display a quadrant pattern which is consistent with the coseismic vertical deformation, suggesting that it is indeed resulting from tectonic processes.  However, the vertical postseismic deformation signal is only apparent in the early postseismic period (figure \ref{fig:NearField2}).  As with the far field deformation, there is a general trend of uplift in the near field one year after the earthquake which we do not consider to be related to postseismic processes.

\section{Postseismic Modeling}\label{sec:Model}

In this paper, we seek to find the mechanisms driving five years of postseismic deformation following the El Mayor-Cucapah earthquake. We consider afterslip and viscoelastic relaxation in the lithosphere and/or asthenosphere as candidate mechanisms.  Poroelastic rebound has also been used to model postseismic deformation \citep[e.g.][]{Jonsson2003}; however, \cite{Gonzalez-ortega2014} suggest that any contribution to postseismic deformation from poroelastic rebound would be negligible.  Additionally, we are considering stations which are sufficiently far away from the main rupture that poroelastic rebound should be insignificant.

We estimate coseismic and time dependent postseismic fault slip, both of which are assumed to occur on the fault geometry estimated by \citet{Wei2011} with some modifications.  Field studies \citep{Fletcher2014} and LIDAR observations \citep{Oskin2012} have revealed a significantly more complicated fault geometry than what was inferred by \citet{Wei2011}, especially within the Sierra Cucapah.  However, we find the relatively simple coseismic fault geometry from \cite{Wei2011} to be sufficient because most of the stations used in this study are sufficiently far from the El Mayor-Cucapah rupture zone that they are insensitive to the details in the fault geometry found by \cite{Fletcher2014} and \cite{Oskin2012}.  The fault geometry used in this study (figure \ref{fig:ContextMap}) consists of the two main fault segments inferred by \cite{Wei2011}, where the northern segment runs through the Sierra Cucapah up to the US-Mexico border and the southern segment is the Indiviso fault which extends down to the Gulf of California.  Additionally, We extend the northern segment by 40 km to the northwest which is motivated by the clustering of aftershocks on the northern tip of the coseismic rupture zone \citep{Hauksson2011,Kroll2013}.  This extended fault segment was also found to be necessary by \cite{Rollins2015} and \cite{Pollitz2012} for describing postseismic deformation.

We consider a variety of rheologic models for the lithosphere and asthenosphere in this study.  The simplest rheologic model is to consider the lithospere and asthenosphere to be effectively elastic and isotropic.  In such case, the rheologic parameters, consisting of the Lame parameters, $\lambda$ and $\mu$, and density, $\rho$, are reasonably well known. The only unknowns are then the distribution of fault slip, which can be easily estimated due to the linearity of the forward problem.  \cite{Rollins2015} found that postseismic deformation following the El Mayor-Cucapah earthquake can be well explained with afterslip on the coseismicly ruptured fault, which has been extended down to the lithosphere-asthenosphere boundary inferred by \cite{Lekic2011}, and they did not require any viscoelastic relaxation to describe the observations.  They did however, find that an unrealistically large amount of afterslip is required on the Indiviso fault, equivalent to a Mw7.2 earthquake.  Since most of the GPS data is located north of the rupture zone, slip on the Indiviso fault segment is not well constrained and it could be acting as a proxy for distributed relaxation in the lower crust or upper mantle.  It is well understood that afterslip at sufficiently great depths can explain postseismic deformation resulting from viscoelastic relaxation \cite{Savage1990}, at least in 2D earthquake models.  In the interest of eliminating null space in our inverse problem,  we assume that afterslip occurs only at seismogenic depths ($<15$ km), and we allow for viscoelastic relaxation in the lower crust (15-30 km) and the upper mantle.  We choose such shallow constraints on afterslip because rate-state friction based models would suggest afterslip to occur at shallow depths to accommodate a coseismic slip deficit \cite{Marone1991}, which is indeed inferred in the coseismic models by \cite{Wei2011a} and (Fialko 2010).  Additionally, as noted by \cite{Rollins2015} the postseismic uplift just north of the rupture zone can be well described with slip at seismogenic depths, while afterslip which is constrained to depths below 15 km predicts subsidence in that region.

Another reason for constraining slip to be shallower than 15 km depth is because we seek to estimate a lower crustal viscosity and the inverse problem becomes particularly ill-posed when trying to simultaneously estimate afterslip and a lithospheric viscosity at the same depths. 
This ill-posedness is illustrated in figure (\ref{fig:LowerCrust}), which shows the displacement resulting from a meter of slip on a fault extending from 15 to 30 km depth as well as the initial velocity resulting from viscoelastic relaxation in the lower crust, which is given a viscosity of $10^{18}$ Pa s.  The horizontal displacement from fault slip is in the opposite direction as the displacements resulting from subsequent viscoelastic relaxation.  This means that displacement resulting from afterslip at lower crustal depths can be cancelled out, at least partially, by a low viscosity lower crust.  We eliminate this null space by allowing only one mechanism in the lower crust, which we choose to be viscoelastic relaxation.  Aside from the issue of non-uniqueness, it is also nonsensical to allow for both brittle and ductile deformation mechanisms at the same depth.

A schematic representation of the viscoelastic rheologic models considered in this study for the lower crust and upper mantle are shown in figure \ref{fig:Rheology}.  We consider Maxwell viscoelasticity, where the steady state viscosity $\eta_M$ is unknown; Zener viscoelasticty, where the transient viscosity $\eta_K$ and transient shear modulus $\mu_K$ are unknown; and Burgers viscoelasticity, where $\eta_M$, $\eta_K$, and $\mu_K$, are all largely unknown. We further discuss these rheologic models and their use in geophysical studies in the discussion.

\subsection*{constraint on effective viscosity}\label{InitialInversion}
For any linear viscoelastic rheology of the lithosphere and asthenosphere, postseismic displacement resulting from time dependent fault slip, $s(\xi,t)$, on a fault $F$ can be described as  
\begin{equation}\label{GeneralForward}
  u(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F s(\xi,\tau) f(t-\tau,x,\xi) d\xi d\tau
\end{equation}
where $g(x,\xi)$ describes the instantaneous elastic displacements at $x$ resulting from a unit of slip at $\xi$, and $f(t,x,\xi)$ describes the time dependent velocity at $x$ resulting from viscoelastic relaxation of stresses induced by slip at $\xi$. $f$ is a function of the properties of the chosen rheologic model, which are generally not well known.  The parameters controlling the instantaneous elastic behaviour of the lithosphere and asthenosphere, $\mu$, $\lambda$, and density $\rho$, are relatively well known and we use the values listed in table (TBD), which are consistent with the properties used in the coseismic model by \cite{Wei2011} and the 1D SCEC background velocity model.  The remaining rheologic parameter, which we will assume to have a depth dependence, are estimated to optimize the predicted fit to observable postseismic displacement.
